---
title: "World Climate PCA"
author: "Sandra Fogg"
date: "1/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = FALSE, message = FALSE, include = FALSE}
# Load Required Packages
library(tidyverse)
library(janitor)
library(naniar)
library(VIM)
library(skimr)
library(ggfortify)
```
World climatic and environmental variables were obtained from Google Earth Engine and compiled by Zander Venter on Kaggle. Mean values for each country included measurements of rainfall levels, temperature extremes, weather conditions, and topography. Principle components analysis (PCA) is utilized below to better understand patterns of climatic variation across the globe. 

### Data Processing
```{r, warning = FALSE, message = FALSE}
## Clean the Data
# Read in the Data
world_climate_df <- read_csv("world_env_vars.csv") 

# Disable Scientific Notation
options(scipen = 999)

# Select the Columns of Interest
# Omit NAs
# Convert Column Headers and Row Strings to Snakecase
# Remove Country Column

climate_clean_df <- world_climate_df %>% 
  select(Country, rain_mean_annual, temp_mean_annual, elevation, tree_canopy_cover, cropland_cover) %>% 
  na.omit() %>% 
  clean_names(case = "snake") %>% 
  mutate(country_l = make_clean_names(country, case = "snake")) %>% 
  select(-country)
```

### Running PCA
```{r, warning = FALSE, message = FALSE}
##Select Indicator Variables
climate_subset <- climate_clean_df %>% 
  select(rain_mean_annual:cropland_cover)

```

```{r, warning = FALSE, message = FALSE, include = FALSE}
## Summarize the Subset
summary(climate_subset) 
```

```{r, warning = FALSE, message = FALSE}
## Run PCA Using `procomp`
climate_pca <- prcomp(climate_subset, scale = TRUE) 
```

## Results
```{r, warning = FALSE, message = FALSE}
# Plot the Variances
plot(climate_pca)

# Clean the biplot
clean_biplot <- autoplot(climate_pca,
                         colour = NA,
                       loadings.label.size = 3,
                       loadings.label.colour = "black",
                       loadings.label.repel = TRUE,
                       loadings.label = TRUE) +
  theme_minimal() +
  labs(title = "Principle Components Analysis of Worldwide Climate Data", vars) 

clean_biplot
```

What are a few main things we can take out of this? What are the main correlations you notice? Are they in line with what you would expect, or is anything surprising? 


A summary (bulletpoints are fine) pointing out some major takeaways from your PCA that can be gleaned from the biplot
You are also welcome to include a photo in the HTML to add some visual interest (not required for Task 2).
